Copulas and their potential for ecology

### 2019-12-18 Paper.Rmd
changed distribution function to cdf

Well-developed approaches do exist, however, and are 
applied widely in other fields, 
that provide a description of the dependence 
between two or more variables that can be mathematically proven 
to be complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
The approaches are based on 
the concept of the *copula*. Copula approaches separate the information of a bivariate 
random variable into two distinct parts: the information in the 
marginal distributions (which says nothing about the dependence between the 
variables), and the remaining information (which is solely about
dependence). We here introduce some simple concepts, based on taking ranks,
that relate to copulas and give a conceptual 
flavor of copula ideas to be introduced 
more formally in the "Background on copulas" section below (section \ref{Background}). Given 
a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the dependence
between $x$ and $y$ can be separated from information contained 
in the marginal distributions by 
considering the bivariate plot of $u_i$ versus $v_i$, where
$u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The rank
of the smallest element in a set is $1$. The $u_i$ and 
$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$; they relate to 
the empirical cumulative distribution function of the $x_i$ and $y_i$,
respectively. We henceforth abbreviate "cumulative distribution function" as "distribution
function" or "cdf". See, for instance,
Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
A-B, respectively. Ranking makes the marginal distributions of 
the component datasets uniform, isolating the 
dependence structure. Dependence structure and marginals can then be studied 
separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
relate to copulas in a way to be described in the "Background on copulas" section below
(section \ref{Background}), after a definition of copulas
is provided. Note that Pearson correlation, even though it is the most commonly
used measure of relationships between variables, is modified by monotonic
tranformation of the component variables (Appendix 
\ref{SM-pedagog1_details}, Fig. \ref{SM-fig_transf_pearson_spearman}),
and therefore reflects not only dependence information, but also
information on the marginals [@Genest2007]. 
Spearman and Kendall correlations, being rank-based, 
are not influenced by monotonic transformations of the component variables. 
They provide information solely on the dependence between those variables.

@@

We give a brief introduction to copulas, focusing, for simplicity, on bivariate 
copulas and on concepts needed
for the rest of the paper. We assume familiarity with the basic 
language of probability distributions and random variables.
See @nelsen2006_copula, @Genest2007, @joe2014_dependence, and
@MaiScherer2017 for general background on copulas, 
and see also @Anderson2018, part of 
which is a short introduction to 
copulas for ecologists. A bivariate *copula* can be defined as a bivariate distribution
function with both margins uniform on $(0,1)$ (@joe2014_dependence, p. 7). 
This will be the
distribution function of a bivariate random variable with uniform marginals, and 
the terminology *copula* is sometimes also applied to this random vector. 
A foundational theorem of
@sklar1959_theorem (see also @MaiScherer2017, p. 16) 
says that if $F$ is the bivariate distribution function
of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$, then there 
exists a copula $C$
such that for all $(x,y)$ in the Euclidean plane, $F(x,y)=C(F_X(x),F_Y(y))$.
The theorem also says $C$ is unique if $F_X$ and $F_Y$ are continuous.
Thus $C$ couples the bivariate distribution function, $F$, with the 
distribution functions of the marginals. 
Finally, Sklar's theorem says that if $D$ is any bivariate copula and 
$G_A$ and $G_B$ are univariate distribution functions, then 
$D(G_A(x),G_B(y))$ is a valid distribution function 
of some bivariate random variable.
The applications of this study fall within the case in which univariate marginal 
distributions have continuous, strictly 
monotonic distribution functions, and this case is simpler. So we 
henceforth make such an assumption.

Sklar's theorem implies that
any random vector can be constructed from a unique copula and marginal 
distribution functions, and, furthermore,
any copula and any univariate distribution functions give rise to a random vector;
so Sklar's theorem provides a correspondence between random vectors
and pairs consisting of a copula and two univariate distribution functions,
thereby making it possible to study a random vector by studying these two
constituents.
Marginals contain no information about the
dependence structure of a random vector, so the copula contains
all such information - it is a complete description of the dependence
between variables.
<!--***-->
The univariate distribution functions associated with a random vector 
$(X,Y)$ are its margins, $F_X$ and $F_Y$, and the 
associated copula is the copula, $C$, guaranteed by Sklar's theorem
(and guaranteed unique, thanks to the assumptions of continuity 
and monotonicity made above). In fact, and making Sklar's correspondence
more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
that $C$ is the bivariate distribution
function of the random variable $(F_X(X),F_Y(Y))$. 
<!--***-->
Conversely, given a copula
$D$ and univariate distribution functions $G_A$ and $G_B$,
Sklar's theorem guarantees that $D(G_A(x),G_B(y))$ is a 
distribution function, so the random variable with this 
distribution function is the random variable corresponding
to $D$, $G_A$ and $G_B$ under Sklar's correspondence. 
Again making Sklar's correspondence
more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
that this random variable is $(G_A^{-1}(U),G_B^{-1}(V))$, 
where $G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
and $(U,V)$ is the random variable with distribution function $D$
(which has uniform marginals).
<!--***-->
If we conflate a copula with the random vector of 
which the copula is the distribution function, then Sklar's correspondence
is between random variables, $(U,V)$, with uniform marginals 
and random variables, $(X,Y)$, with arbitrary continuous, strictly monotonic
marginals. The correspondence is simply via the application of the
univariate distribution functions, or their inverses: $U=F_X(X)$,
$V=F_Y(Y)$, $X=F^{-1}_X(U)$, $Y=F^{-1}_Y(V)$.
<!--***-->
Application of the 
distribution functions of the component random variables translates
between the uniform-marginal context, $(U,V)$, which isolates the 
dependence information, and the context of the original random variable,
$(X,Y)$.

@@

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family, which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of association
between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cummulative dstribution function (cdf) of a (univariate) standard 
normal and $\Phi_{2,p}$ is the joint cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
directly informative about copula properties as probability density function (pdf) 
pictures; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate normal copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails. 
The Clayton copula family,
of which we have already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). Higher values of the parameter,
$p$ produce copulas from the Clayton family with higher Kendall or Spearman correlation. The survival
Clayton family is the symmetric opposite of the Clayton family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
information on these and other copula families.

@@

Plots using normalized ranks were proposed in the Introduction as being conceptually
similar to copulas. We here explain the connection. As stated above, the copula associated
with a random variable $(X,Y)$ is the joint distribution function of the random variable
$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$, from $(X,Y)$, let $\hat{F}_X$ and
$\hat{F}_Y$ be the empirical cdfs associated with the $x_i$ and $y_i$, respectively.
These are step functions of $x$ and $y$, respectively, that start at $0$ for low values of 
$x$ and $y$ and jump by $1/n$
at each of the data points. Therefore $\hat{F}_X(x_j)$ is the rank of $x_j$ in
the set $\{x_i : i=1,\ldots,n\}$, dividided by $n$, 
and analogously $\hat{F}_Y(y_j)$ is the rank of $y_j$ in the set 
$\{y_i : i=1,\dots,n\}$, divided
by $n$. For large $n$, $\hat{F}_X(x_i)$ and $\hat{F}_Y(y_i)$ 
approximate the cdfs $F_X$ and $F_Y$, respectively. 
But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
$n/(n+1)$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
$n/(n+1)$. So the normalized rank in turn approximates the empirical cdf and therefore 
the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
regarded as approximate samples from the random variable $(F_X(X),F_Y(Y))$, 
which is the random variable associated with the copula of $(X,Y)$.
The scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula 
structure (see section \ref{Methods} below).

@@

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described in more detail elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail dependence 
of the families, and provide pictorial descriptions (Figs 
\ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
The pictorial decriptions can also be used to aid quick comparisons of copula
families when alternative families are being considered for future applications. 
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 families. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters. See 
Table \ref{tab_copula_info} for a summary of some properties of each family.
<!--***DAN: Shyamolina, please make a table, and replace the X above with 
a citation to that table. Please place the table chunk in such a way that 
this citation becomes table 2, since prior to this point the only main text
table that has been cited is table 1, and we want the table citations to 
come in order. The columns of the table should include: 1) copula family name,
2) number of parameters for the family; 3) info on tail dependence that is
in the above para; 4) a reference to the figure (in the main text or sup mat)
where there is a picture of this family. Anything else? There should be a 
row for each of our 16 copula families.-->

@@

<!--adding plots showing influences of non-normal copula on spatial Taylors law-->
\begin{figure}[!h]
\textbf{ \hspace{2cm} (A) \hspace{2.92 cm} (B) \hspace{2.92 cm} (C) \hspace{2.92 cm} (D)} \\
\vspace{-0.8 cm} 
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
\end{center}
\textbf{ \hspace{1 cm} (E) \hspace{1.4 cm} (F) \hspace{1.2 cm} (G) \hspace{1.2 cm} (H) \hspace{1.4 cm} (I) \hspace{1.4 cm} (J) \hspace{1.2 cm} (K)}\\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
locations. Results from all $1000$ 
simulations were plotted on the same axes for each panel, but two example simulations are 
shown in solid and void black circles to help assess to what extent variation on the plots was between 
simulations or within simulations. For each simulation, we tested the linearity 
(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
and quantified the root mean squared error of points from the linear regression line (G).
We also calculated the linear regression intercept (H) and slope (I), the quadratic
term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
(J), and the curvature of that quadratic regression (K). Distributions of values 
across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
though results using Kendall-preserving surrogates were similar. See section \ref{Consequences}
and Appendices \ref{SM-Consequence_TL_simulation} and 
\ref{SM-sect:copsurrognd} for details.\label{fig_spTL}}
\end{center}
\end{figure}

@@

@@

<!--additional copula methods-->
Copula methods are numerous, and go well beyond the cases we have considered
in this introductory paper. We hope our work helps inspire applications of copulas
in ecology both of and 
beyond the specific tools we have employed. For instance, multivariate copulas are useful for 
studying relationships between multiple interacting variables, and have been used in finance
to model tail risk for portfolio optimization problems [@Czado2019; @Joe2011]. 
Analogous ecological applications
exist that may be amenable to the same approach, e.g., ecosystem functioning variables
such as total primary productivity or carbon flux are the sum of the contributions of
multiple species in the same way that the value of an investment portfolio is the sum 
of the values of its constituent assets. Ecologists may likewise be interested in
managing the risk that an ecosystem functioning variable will take an extreme value. 
As another example, we assumed for this paper that univariate marginal 
distributions have continuous, strictly 
monotonic distribution functions. This was for simplicity and because the simpler
approach was sufficient for our research questions. But count data in ecology have been analyzed
with approaches not making this assumption [@Anderson2018].

>>>>

Well-developed approaches do exist, however, and are 
applied widely in other fields, 
that provide a description of the dependence 
between two or more variables that can be mathematically proven 
to be complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
The approaches are based on 
the concept of the *copula*. Copula approaches separate the information of a bivariate 
random variable into two distinct parts: the information in the 
marginal distributions (which says nothing about the dependence between the 
variables), and the remaining information (which is solely about
dependence). We here introduce some simple concepts, based on taking ranks,
that relate to copulas and give a conceptual 
flavor of copula ideas to be introduced 
more formally in the "Background on copulas" section below (section \ref{Background}). Given 
a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the dependence
between $x$ and $y$ can be separated from information contained 
in the marginal distributions by 
considering the bivariate plot of $u_i$ versus $v_i$, where
$u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The rank
of the smallest element in a set is $1$. The $u_i$ and 
$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$; they relate to 
the empirical cumulative distribution function of the $x_i$ and $y_i$,
respectively. We henceforth abbreviate "cumulative distribution function" as "cdf". 
See, for instance,
Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
A-B, respectively. Ranking makes the marginal distributions of 
the component datasets uniform, isolating the 
dependence structure. Dependence structure and marginals can then be studied 
separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
relate to copulas in a way to be described in the "Background on copulas" section below
(section \ref{Background}), after a definition of copulas
is provided. Note that Pearson correlation, even though it is the most commonly
used measure of relationships between variables, is modified by monotonic
tranformation of the component variables (Appendix 
\ref{SM-pedagog1_details}, Fig. \ref{SM-fig_transf_pearson_spearman}),
and therefore reflects not only dependence information, but also
information on the marginals [@Genest2007]. 
Spearman and Kendall correlations, being rank-based, 
are not influenced by monotonic transformations of the component variables. 
They provide information solely on the dependence between those variables.

@@

We give a brief introduction to copulas, focusing, for simplicity, on bivariate 
copulas and on concepts needed
for the rest of the paper. We assume familiarity with the basic 
language of probability distributions and random variables.
See @nelsen2006_copula, @Genest2007, @joe2014_dependence, and
@MaiScherer2017 for general background on copulas, 
and see also @Anderson2018, part of 
which is a short introduction to 
copulas for ecologists. A bivariate *copula* can be defined as a bivariate cdf 
with both margins uniform on $(0,1)$ (@joe2014_dependence, p. 7). 
This will be the 
cdf of a bivariate random variable with uniform marginals, and 
the terminology *copula* is sometimes also applied to this random vector. 
A foundational theorem of
@sklar1959_theorem (see also @MaiScherer2017, p. 16) 
says that if $F$ is the cdf
of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$, then there 
exists a copula $C$
such that for all $(x,y)$ in the Euclidean plane, $F(x,y)=C(F_X(x),F_Y(y))$.
The theorem also says $C$ is unique if $F_X$ and $F_Y$ are continuous.
Thus $C$ couples the bivariate cdf, $F$, with the 
cdfs of the marginals. 
Finally, Sklar's theorem says that if $D$ is any bivariate copula and 
$G_A$ and $G_B$ are univariate cdfs, then 
$D(G_A(x),G_B(y))$ is a valid cdf 
of some bivariate random variable.
The applications of this study fall within the case in which univariate marginal 
distributions have continuous, strictly 
monotonic cdfs, and this case is simpler. So we 
henceforth make such an assumption.

Sklar's theorem implies that
any random vector can be constructed from a unique copula and marginal 
cdfs, and, furthermore,
any copula and any univariate cdfs give rise to a random vector;
so Sklar's theorem provides a correspondence between random vectors
and pairs consisting of a copula and two univariate cdfs,
thereby making it possible to study a random vector by studying these two
constituents.
Marginals contain no information about the
dependence structure of a random vector, so the copula contains
all such information - it is a complete description of the dependence
between variables.
<!--***-->
The univariate cdfs associated with a random vector 
$(X,Y)$ are its margins, $F_X$ and $F_Y$, and the 
associated copula is the copula, $C$, guaranteed by Sklar's theorem
(and guaranteed unique, thanks to the assumptions of continuity 
and monotonicity made above). In fact, and making Sklar's correspondence
more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
that $C$ is the bivariate distribution
function of the random variable $(F_X(X),F_Y(Y))$. 
<!--***-->
Conversely, given a copula
$D$ and univariate cdfs $G_A$ and $G_B$,
Sklar's theorem guarantees that $D(G_A(x),G_B(y))$ is a 
cdf, so the random variable with this 
cdf is the random variable corresponding
to $D$, $G_A$ and $G_B$ under Sklar's correspondence. 
Again making Sklar's correspondence
more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
that this random variable is $(G_A^{-1}(U),G_B^{-1}(V))$, 
where $G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
and $(U,V)$ is the random variable with cdf $D$
(which has uniform marginals).
If we conflate a copula with the random vector of 
which the copula is the cdf, then Sklar's correspondence
is between random variables, $(U,V)$, with uniform marginals 
and random variables, $(X,Y)$, with arbitrary continuous, strictly monotonic
marginals. The correspondence is simply via the application of the
univariate cdfs, or their inverses: $U=F_X(X)$,
$V=F_Y(Y)$, $X=F^{-1}_X(U)$, $Y=F^{-1}_Y(V)$.
Application of the 
cdfs of the component random variables translates
between the uniform-marginal context, $(U,V)$, which isolates the 
dependence information, and the context of the original random variable,
$(X,Y)$.

@@

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family, which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of association
between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cummulative dstribution function (cdf) of a (univariate) standard 
normal and $\Phi_{2,p}$ is the cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. Formulas for the normal copula and all the 
other copula families we use were provided by @Brechmann2013. 
However, formulas for copulas are often not as 
directly informative about copula properties as probability density function (pdf) 
pictures; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate normal copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails. 
The Clayton copula family,
of which we have already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). Higher values of the parameter,
$p$ produce copulas from the Clayton family with higher Kendall or Spearman correlation. The survival
Clayton family is the symmetric opposite of the Clayton family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula, @joe2014_dependence and 
@Brechmann2013 for further
information on these and other copula families.

@@

Plots using normalized ranks were proposed in the Introduction as being conceptually
similar to copulas. We here explain the connection. As stated above, the copula associated
with a random variable $(X,Y)$ is the cdf of the random variable
$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$, from $(X,Y)$, let $\hat{F}_X$ and
$\hat{F}_Y$ be the empirical cdfs associated with the $x_i$ and $y_i$, respectively.
These are step functions of $x$ and $y$, respectively, that start at $0$ for low values of 
$x$ and $y$ and jump by $1/n$
at each of the data points. Therefore $\hat{F}_X(x_j)$ is the rank of $x_j$ in
the set $\{x_i : i=1,\ldots,n\}$, dividided by $n$, 
and analogously $\hat{F}_Y(y_j)$ is the rank of $y_j$ in the set 
$\{y_i : i=1,\dots,n\}$, divided
by $n$. For large $n$, $\hat{F}_X(x_i)$ and $\hat{F}_Y(y_i)$ 
approximate the cdfs $F_X$ and $F_Y$, respectively. 
But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
$n/(n+1)$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
$n/(n+1)$. So the normalized rank in turn approximates the empirical cdf and therefore 
the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
regarded as approximate samples from the random variable $(F_X(X),F_Y(Y))$, 
which is the random variable associated with the copula of $(X,Y)$.
The scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula 
structure (see section \ref{Methods} below).

@@

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described in more detail elsewhere [@Joe1997; @Brechmann2013]; so 
we only briefly identify each family, say a few words about tail dependence 
of the families, and provide pictorial descriptions (Figs 
\ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
The pictorial decriptions can also be used to aid quick comparisons of copula
families when alternative families are being considered for future applications. 
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 families. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 and @Brechmann2013 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters. See 
Table \ref{tab_copula_info} for a summary of some properties of each family.

@@

<!--adding plots showing influences of non-normal copula on spatial Taylors law-->
\begin{figure}[!h]
\textbf{ \hspace{2cm} (A) \hspace{2.92 cm} (B) \hspace{2.92 cm} (C) \hspace{2.92 cm} (D)} \\
\vspace{-0.8 cm} 
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
\end{center}
\textbf{ \hspace{1 cm} (E) \hspace{1.4 cm} (F) \hspace{1.2 cm} (G) \hspace{1.2 cm} (H) \hspace{1.4 cm} (I) \hspace{1.4 cm} (J) \hspace{1.2 cm} (K)}\\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
locations. Results from all $1000$ 
simulations were plotted on the same axes for each panel, but two example simulations are 
shown in solid and hollow black circles to help assess to what extent variation on the plots was between 
simulations or within simulations. For each simulation, we tested the linearity 
(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
and quantified the root mean squared error of points from the linear regression line (G).
We also calculated the linear regression intercept (H) and slope (I), the quadratic
term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
(J), and the curvature of that quadratic regression (K). Distributions of values 
across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
though results using Kendall-preserving surrogates were similar. See section \ref{Consequences}
and Appendices \ref{SM-Consequence_TL_simulation} and 
\ref{SM-sect:copsurrognd} for details.\label{fig_spTL}}
\end{center}
\end{figure}

@@

<!--referee-requested para comparing to regression with heteroskedastic error term-->
Non-copula methods may be useful under some circumstances for exploring
asymmetries of tail association, but copula approaches have theoretical
and practical advantages. It was suggested to us that asymmetries of
tail association can also be explored using, for instance, a regression 
model with heteroskedastic error term, $y=mx+b+\epsilon(x)$. Whereas this 
approach seems capable of yielding insight for some datasets, it implicitly assumes
a particular causal relationship between the variable, an untenable assumption for our
synchrony applications and many other applications. Furthermore, no equivalent
of Sklar's theorem exists under this approach, so information on marginal 
distributions cannot be separated from association information. Sklar's theorem
means that the copula associated with two variables contains all the information
about the association. Such statements of 
mathematical completeness will not be avilable for others methods.

@@

<!--additional copula methods-->
Copula methods are numerous, and go well beyond the cases we have considered
in this introductory paper. We hope our work helps inspire applications of copulas
in ecology both of and 
beyond the specific tools we have employed. For instance, multivariate copulas are useful for 
studying relationships between multiple interacting variables, and have been used in finance
to model tail risk for portfolio optimization problems [@Czado2019; @Joe2011]. 
Analogous ecological applications
exist that may be amenable to the same approach, e.g., ecosystem functioning variables
such as total primary productivity or carbon flux are the sum of the contributions of
multiple species in the same way that the value of an investment portfolio is the sum 
of the values of its constituent assets. Ecologists may likewise be interested in
managing the risk that an ecosystem functioning variable will take an extreme value. 
As another example, we assumed for this paper that univariate marginal 
distributions have continuous, strictly 
monotonic cdfs. This was for simplicity and because the simpler
approach was sufficient for our research questions. But count data in ecology have been analyzed
with approaches not making this assumption [@Anderson2018].
